# Communicable, Maternal, Neonatal, and Nutritional Disease Burden (CMNN)

```{r include = F}
knitr::opts_chunk$set(echo = FALSE)
```


```{r message=FALSE}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here)

# Import the communicable diseases data
data_cmnn <- read_csv(here("data", "burden-of-disease-cmnn.csv"))
```

## Table of Estimates for CMNN Burden Over Time

```{r}
# Here render a table for the DALY burden over time for the three countries 
# You should pivot the data to show each country in a separate column.
# Use kable() from the knitr package to render the table.

cmnn_burden_table <- 
  data_cmnn %>% 
  filter(Entity %in% c("Burkina Faso", "Madagascar", "New Zealand")) %>%
  arrange(Year) %>% 
  rename(DALY = `DALYs (Disability-Adjusted Life Years) - Communicable, maternal, neonatal, and nutritional diseases - Sex: Both - Age: Age-standardized (Rate)`) %>% 
  select(!Code) %>% 
  pivot_wider(values_from = DALY, 
              names_from = Entity)

  knitr::kable(cmnn_burden_table)


```

## Plot Showing Trends in CMNN Burden Over Time

```{r}
# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.

cmnn_burden_graph <-
  
ggplot(cmnn_burden_table, 
       aes(x = Year, y = `Burkina Faso`, color = "Burkina Faso")) +
  geom_line() +
  
  geom_line(aes(y = `Madagascar`, color = "Madagascar")) +
  
  geom_line(aes(y = `New Zealand`, color = "New Zealand")) +
  
  labs(title = "DALYs Burden Over Time for communicable, maternal, neonatal, and nutritional diseases",
       x = "Year",
       y = "DALYs",
       color = "Country") +
  
  scale_color_manual(values = c("Burkina Faso" = "blue", "Madagascar" = "green", "New Zealand" = "red")) +
  scale_y_log10() +
  
  theme_minimal()


cmnn_burden_graph

```

## Summary of CMNN Burden Findings

Disease burden for communicable, maternal, neonatal and nutritional diseases seems to steadily, although not rapidly decrease in all three countries. The decrease is less pronounced in New Zealand, where we might assume access to higher quality care and medicine has been more constant over the years. Access to such care in the other two countries seems to have had a greater impact, with a very similar rate of decline in both.


