# Non-Communicable Disease Burden (NCD)

```{r include = F}
knitr::opts_chunk$set(echo = FALSE)
```

```{r include=FALSE}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here)

# Import the non-communicable diseases data
data_ncd <- read_csv(here("data", "burden-of-disease-ncd.csv"))
```

## Table of Estimates for NCD Burden Over Time

```{r}
# Here render a table for the DALY burden over time for the three countries 
# You should pivot the data to show each country in a separate column.
# Use kable() from the knitr package to render the table.

ncd_burden_table <- 
  data_ncd %>% 
  filter(Entity %in% c("Burkina Faso", "Madagascar", "New Zealand")) %>%
  arrange(Year) %>% 
  rename(DALY = `DALYs (Disability-Adjusted Life Years) - Non-communicable diseases - Sex: Both - Age: Age-standardized (Rate)`) %>% 
  select(!Code) %>% 
  pivot_wider(values_from = DALY, 
              names_from = Entity)

  knitr::kable(ncd_burden_table)


```

## Plot Showing Trends in NCD Burden Over Time 

```{r}
# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.

ncd_burden_graph <-
  
ggplot(ncd_burden_table, 
       aes(x = Year, y = `Burkina Faso`, color = "Burkina Faso")) +
  geom_line() +
  
  geom_line(aes(y = `Madagascar`, color = "Madagascar")) +
  
  geom_line(aes(y = `New Zealand`, color = "New Zealand")) +
  
  labs(title = "DALYs Burden Over Time for Non communicable diseases",
       x = "Year",
       y = "DALYs",
       color = "Country") +
  
  scale_color_manual(values = c("Burkina Faso" = "blue", "Madagascar" = "green", "New Zealand" = "red")) +
  scale_y_log10() +
  
  theme_minimal()


ncd_burden_graph


```

## Summary of NCD Burden Findings

Although the findings on other diseases burden shows a general downward trend in all 3 countries, the graph above. for non-communicable diseases, seems to indicate that some countries are seeing an uptick since 2010. Burkina Faso and New Zealand both see a slight increase, although slightly more pronounced in the former, which ends up overtaking Madagascar in 2015. Madagascar seems to see a more steady decline over the years, although also less pronounced than for other forms of disease burden.


