# Overall Disease Burden
```{r include = F}
knitr::opts_chunk$set(echo = FALSE)
```


```{r include = FALSE}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here, esquisse)

# Import the overall disease data
data_overall <- read_csv(here("data", "burden-of-disease-all-causes.csv"))
```

## Table of Estimates for Overall Disease Burden Over Time

```{r echo=}
# Here render a table for the DALY burden over time for the three countries 
# You should pivot the data to show each country in a separate column.
# Use kable() from the knitr package to render the table.

daly_burden_table <- 
  data_overall %>% 
  filter(Entity %in% c("Burkina Faso", "Madagascar", "New Zealand")) %>%
  arrange(Year) %>% 
  rename(DALY = `DALYs (Disability-Adjusted Life Years) - All causes - Sex: Both - Age: Age-standardized (Rate)`) %>% 
  select(!Code) %>% 
  pivot_wider(values_from = DALY, 
              names_from = Entity)

  knitr::kable(daly_burden_table)
```

## Plot Showing Trends in Overall Disease Burden Over Time

```{r echo=FALSE}
# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.

daly_burden_graph <-
  
ggplot(daly_burden_table, 
       mapping = aes(x = Year, y = `Burkina Faso`, color = "Burkina Faso")) +
  geom_line() +
  
  geom_line(aes(y = `Madagascar`, color = "Madagascar")) +
  
  geom_line(aes(y = `New Zealand`, color = "New Zealand")) +
  
  labs(title = "DALYs Burden Over Time",
       x = "Year",
       y = "DALYs",
       color = "Country") +
  
  scale_color_manual(values = c("Burkina Faso" = "blue", "Madagascar" = "green", "New Zealand" = "red")) +
  scale_y_log10() +
  
  theme_minimal()


daly_burden_graph
```

## Summary of Overall Disease Burden Findings

The graph above shows decreasing Disability-Adjusted Life Years burden in all three countries between 1990 and 2003.The decrease is slightly sharper in the case of Burkina Faso, although we can observe a short spike in the mid 1990's. New Zealand's decrease is steadier and less pronounced. This visualization suggests progress in access to health care and general advances in medicine had a more significant impact in less industrialized countries than in a country like New Zealand, where such access and quality of health services have presumably always been greater.
